Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-10-09 GWLDCII.tex
grammar and spelling readthrough of sections 1 through 4

\begin{abstract}
  Wilson loop diagrams are an important tool in studying scattering amplitudes of SYM $N=4$ theory and are known by previous work to be associated to positroids. In this paper we study the structure of the associated positroids, as well as the structure of the denominator of integral defined by each diagram. In particular, we give an algorithm to derive the Grassman necklace of the assoicated positroid directly from the Wilson loop diagram. We also give a recursive proof that the dimension of these cells is thrice the number of propagators in the diagram. We also show that the denominator in the integral is the radical of the product of minors defined by the elements of the Grassmann necklace.
\end{abstract}
\section{Introduction}

@@

In recent years, there has been significant interest in understanding the geometry and combinatorics underlying the field theory SYM $N=4$ \cite{wilsonloop, Arkani-Hamed:2013jha, Amplituhedronsquared, galashinlam, AmplituhedronDecomposition}. This started with the observation that BCFW diagrams, which represent the on shell interactions of this theory, correspond to plabic graphs \cite{GrassmannAmplitudebook}. Plabic graphs were introduced by Postnikov in \cite{Postnikov}, and provide one of the many equivalent ways to classify the positroid cells of the positive real Grassmannian $\Gr(k,n)$. This led to the representation of the on shell integrals associated to this theory in terms of a geometric space (the Amplituhedron) embedded in $\Gr(k,n)$ \cite{Arkani-Hamed:2013kca,Arkani-Hamed:2013jha}. Moreover, integrals associated to the BCFW diagrams define a volume form on the Amplituhedron. This both gives a novel and elegant interpretation of scattering amplitudes for this theory, as well as a means of greatly simplifying the calculations involved in traditional Feynman diagram approaches. Since then, there has been significant work to understand the volume form and its physical implications \cite{UnwindingAmplituhedron,galashinlam}.

Meanwhile, a different body of work has studied SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv,Adamo:2012xe,  Boels:2007qn, Bullimore:2010pj}. These integrals are calculated via holomorphic Wilson loops. As with the Amplituhedron, each Wilson loop represents the sum of a family of $N^kMHV$ Feynman diagrams, and adding up the corresponding sums of Feynman integrals associates a volume form to a convex space different from the Amplituhedron \cite{Hodges2013}. Agarwala and Marin-Amat uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that can be realized as an element of $\Gr(k,n)$ \cite{wilsonloop}. Other work has shown a connection between the volume filled by Wilson loop diagrams and the Amplituhedron \cite{casestudy, Amplituhedronsquared}. Specifically, Marcott shows that the space parametrized by the Wilson loop diagrams is  exactly the correct positroid cell in $\Gr(k,n)$, up to a set of measure $0$ \cite{WLDdim}.

@@

This paper continues this work by identifying which positroid cells are associated to which Wilson loop diagram: specifically, we give an algorithm to read off the Grassmann necklace of the associated positroid cell from the Wilson loop diagram. We also show that the positroid cells associated to Wilson loop diagrams are all $3k$-dimensional; by contrast, the BCFW cells are all $4k$-dimensional \cite{Arkani-Hamed:2013jha}. Again, this is not too alarming, as it was shown in \cite{non-orientable} that the actual geometric space defined by the Wilson loop diagrams is a non-orientable $k$-dimensional bundle over this $3k$ dimensional manifold in $\Gr(k,n)$.

@@

In Section~\ref{sec GN algorithm} we state an algorithm for constructing the Grassmann necklace of the positroid associated to a Wilson loop diagram (Algorithm~\ref{alg:put GN on WLD}) and prove its correctness (Theorem~\ref{res:alg gives GN}). While it was already possible to construct the Grassmann necklace of a Wilson loop diagram via existing bijections between some of the various combinatorial objects that index the positroid cells, this process was convoluted and involved multiple steps; our algorithm allows us to track directly how each propagator in the Wilson loop diagram contributes to each term in the Grassmann necklace. Furthermore, the above algorithm allows us to understand the positroids associated to Wilson loop better. In particular, Corollary \ref{no coloops} shows that these positroids have no coloops, and gives an exact characterization of their loops. Many of the technical lemmas required to prove Theorem~\ref{res:alg gives GN} continue to be important in subsequent sections: in particular we have Lemma~\ref{lem sian}, which identifies certain configurations of propagators that must appear in any admissible Wilson loop diagram, and Lemma~\ref{vertex cyclic int lem}, which shows that the pattern of values that a propagator contributes to each term of the Grassmann necklace is both simple and predictable.

@@

\subsection{Wilson Loop diagrams}\label{section WLD background}

@@

A Wilson loop diagram is depicted as a circle, with the vertices $V$ arranged along the edge of the circle and listed counterclockwise in their cyclic order. The arc between two consecutive vertices is referred to as an {\bf edge} of the diagram, and the $i$th edge of $W$ is the edge from vertex $i$ to its successor. Each propagator $(i,j) \in \cP$ is denoted by a wavy line inside the circle joining edge $i$ and edge $j$.

@@

We also impose the convention that all vertices should always be interpreted in terms of the cyclic order on $V$.  So, for any $i\in V$ we will use $i+1$ to denote the successor of $i$ in $V$.   In particular, if $V = [n]$ then $(n,n+3)$ means the propagator connecting edges $3$ and $n$. This viewpoint is beneficial for us since many of our proofs rely on restricting our attention to a region of the diagram bounded by a given propagator, or to short propagators such as $(i,i+2)$, and these behave the same regardless of whether we cross the ``first'' vertex or not; however, we emphasize that this is different to the standard convention in the physics literature, where propagators are written $(i,j)$ with $1 \leq i<j \leq n$.

@@

In particular, we will see in later sections that all four vertices supporting a propagator are important in understanding the contribution of that propagator to the objects represented by the diagram. This motivates the following definition of a subdiagram: any subset $Q$ of the propagators and any subset $U$ of the vertices of $W$ {\em containing the support of $Q$}.

@@

\begin{dfn}\label{admisdfn}
A Wilson loop diagram $W = (\cP, V)$ is called {\bf admissible} if:
\begin{enumerate}
\item $|V| \geq |\cP| + 4$.
\item There does not exist a non-empty set of propagators $Q \subseteq \cP$ such that $|V(Q)| < |Q| + 3$.
\item There does not exist a pair of crossing propagators: that is, there does not exist $\{(i,j),(k,l)\} \subseteq \cP$ such that $i < k < j < l$.
\end{enumerate}
 \end{dfn}

@@

These diagrams correspond to (tree level) amplitudes of SYM $N=4$ theory in twistor space. As such, a diagram associates a real value to $n$ particles, one for each vertex. Each particle is represented as a section of a $|\cP|$-vector bundle over twistor space, projected onto a real subspace in a process called bosonization \cite{Arkani-Hamed:2013jha}. In other words, each Wilson loop diagram defines a functional on external particles. This is represented as an integral $I(W)$, which is the equivalent of a Feynman integral in this setting. This integral is comprised of two main components, and interaction matrix, $C(W)$ and a volume form, $\Omega(W)$. These are described below.

A $|\cP| \times n$ matrix $C(W)$ represents the (tree level) particle interactions. Each row of $C(W)$ corresponds to an internal propagator, each column to an external particle. $C(W)$ is defined as follows:

@@

The matrix $C(W)$ defines a positroid cell associated to $W$ in the positive Grassmannian $\Gr(n, |\cP|)$, which we denote by $\Sigma(W)$ \cite{wilsonloop}. Marcott \cite{WLDdim} shows that  the matrix $C(W)$ parametrizes a space whose close is exactly the closure of $\Sigma(W)$. That is, the space defined by $C(W)$ and the cell $\Sigma(W)$ agree up to a set of measure $0$. Most of the work in this paper, and in \cite{generalcombinatoricsI} focuses on characterizing the positroid cell associated to each Wilson loop diagram $W$.

Note that since there is no ordering on the propagators of $W$, $C(W)$ is only defined up to rearrangement of the rows. When needed, we will impose various orders on these rows. This will not matter, as it does not affect $\Sigma(W)$.

@@

In particular, this implies that any singularities in the integrals $I(W)$ must cancel out in the sum, in particular, all poles arising from $R(W)$ must cancel. Specifically, these poles must cancel on the boundaries of the positroid cells $\Sigma(W)$, analogous to the analysis in \cite{Arkani-Hamed:2013jha}. This has been checked for particular classes of examples (e.g. \cite{casestudy, Amplituhedronsquared, HeslopStewart}) but has not been proven in general.

With this motivation in mind, in this paper we study the denominator $R(W)$ of the volume form $\Omega(W)$, which we define next.

\begin{dfn}\label{def R(W)}
Let $W = (\cP,V)$ be an admissible Wilson loop diagram, and fix an edge $e$. Let $\{p_1 \ldots p_r\}$ be the propagators with one end lying on edge $e$ under a counterclockwise ordering. That is, $p_1$ is closest to the vertex $e$, $p_i$ is positioned counterclockwise to $p_{i+1}$ and $p_r$ is the clockwise most propagator on this edge (closest to vertex $e+1$). Define the polynomial
\[ R_e =  x_{p_1,e+1} \prod_{j= 1}^{r-1} \left((x_{p_j,e} x_{p_{j+1},e+1} - x_{p_{j+1},e} x_{p_{j},e+1} ) \right) x_{p_r,e}\;\]
to be the component of $R(W)$ associated to edge $e$. Note that if $r = 1$ (i.e. there is exactly one propagator lying on edge $e$) then this expression simplifies to $R_e = x_{p,e} x_{p,e+1}$. If $r=0$, set $R_e = 1$. The denominator $R(W)$ is defined to be the product of all of the $R_e$, i.e. 
\[R(W) = \prod_{e \in V} R_e.\]
\end{dfn}

In Section \ref{sec poles}, we show that the denominator $R(W)$ is determined by the positroid cell $\Sigma(W)$ and the structure of $W$, as encapsulated by $C(W)$. In fact, it can be read directly off of $C(W)$, in terms of a characterization of $\Sigma(W)$ called the Grassmann necklace (see Definition~\ref{def:grassmann necklace}) as shown in Algorithm \ref{alg:put GN on WLD}. We also give an algorithm to read $R(W)$ directly from the Wilson loop diagram, paving the way for a general study of these poles in future.

@@

Let $W = (\cP,[n])$ be an admissible Wilson loop diagram with $k$ propagators, and $C(W)$ the associated matrix defined in equation~\eqref{C(W) dfn}. Let $M(W)$ be the matroid realized by $C(W)$, i.e. the matroid with ground set $[n]$ whose independent sets are exactly the sets $V \subseteq [n]$ such that the columns of $C(W)$ indexed by $V$ are linearly independent.

@@

  Alternatively we can simply think about the condition of linear independence in the matrix $C(W)$, the corollary can also be proved directly by linear algebra and the definition of $C(W)$ as follows.

  Because the nonzero entries of $C(W)$ are independent indeterminants, $J$ is an independent set if and only if there is some choice of $|J|$ nonzero entries of $C(W)$ one in each row associated to an element of $J$ and each in different columns. Each entry in $C(W)$ identifies a propagator by the row of the entry and a vertex by the column of the entry.  The entry is nonzero if and only if the propagator is supported on that vertex.  Consequently, a choice of $|J|$ nonzero entries of $C(W)$ one in each row associated to an element of $J$ and each in different columns is equivalent to an assignment of the propagators of $J$ to supporting vertices so that no two are assigned to the same vertex.  Such an assignment of the propagators of $J$ to supporting vertices is exactly a map $f$ as described in the statment, hence proving the result.
\end{proof}

@@

\begin{rmk}
While this paper primarily approaches proofs from a combinatorial point of view, readers with a background in matroid theory are encouraged to keep the matroidal viewpoint in mind throughout; we expect that the two approaches will be complementary and will build upon each other. See \cite{generalcombinatoricsI} for some related results concerning the positroids of Wilson loop diagrams obtained by viewing them primarily as matroids.
\end{rmk}

\section{Wilson Loop diagrams and their Grassmann necklaces}\label{sec GN algorithm}

In this section we will give an algorithm (Algorithm~\ref{alg:put GN on WLD}) to go directly from a Wilson loop diagram to the Grassmann necklace of its positroid. This is a useful result in and of itself, as it greatly simplifies the process of identifying the positroid associated to a given Wilson loop diagram. Furthemore, we use it to characterize loops and coloops in the positroid associated to a Wilson loop diagram (Corrollary \ref{no coloops}). It allows us to see how the Le diagram assoicated to a Wilson loop diagram changes as one adds propagators in certain positions (Theorem \ref{thm dim}). Indeed, the techniques we develop in this section to verify that our algorithm produces the correct Grassmann necklace will also be used repeatedly in later sections to calculate the dimension of the associated positroid cell (section \ref{sec dim}) and to study the structure of the poles of the associated integral $I(W)$ (section \ref{sec poles}).

One of the key insights is that each element of the Grassmann necklace can be viewed as a function from the propagators of the diagram to the vertices in the Grassman necklace element.  This is captured in Definition~\ref{def I_i as a function}, and used throughout the paper.

@@

Before we can describe the algorithm for extracting the Grassmann necklace of $M(W)$ from the Wilson loop diagram $W$, we require some initial results about the behavior of propagators in admissible WLD.

@@

Let $p$ be the propatagor $p = (1, 4)$. Then $\cP_{in}(p, <_4) = \{(1,4), (5, 8), (5, 7) \}$ and $\cP_{out}(p, <_4) = \{(2,4) \}$
\end{eg}

@@

\begin{rmk}\label{rem:props of length 2 and 3} 
  The following observations about configurations of propagators of short length in a weakly admissibly Wilson loop diagram $W$ are easily verified:
 \begin{enumerate}
\item If $p = (i,i+3)$ is a propagator of length 3, then the middle vertex $i+2$ supports at most one propagator.
\item If every vertex in $W$ supports at least one propagator, then $W$ admits at least one propagator of length 2.  Furthermore, by the same reasoning, if $p=(i,j)$ is a propagator of $W$ and every vertex in $[i,j]$ supports at least one propagator then there is at least one propagator of length $2$ in $\mathcal{P}_{in}(p, <_i)$.
\end{enumerate}
\end{rmk}

\begin{figure}[h]
\[
\begin{tikzpicture}[baseline=(current bounding box.north)]
  \begin{scope}
  \drawWLDfragment[7]{3}{0.4} 
  %\drawnumberspartial % use this to debug
  \centerarc[linehighlight](0,0)(\zero + 1.43*\step:\zero+6.6*\step:\radius)
  \newnode{1}{}
  \newnode{3.55}{}
  \newnode{4.45}{}
  \newnode{7}{}
    \begin{scope}
    \clipcenterarc(0,0)(\startpoint:\endpoint:\radius)
  \newpropbend{1.3}{6.7}{20}
  \newpropbend{2.9}{5.1}{45}
  \end{scope}
  \node[align=center] at (\zero + 4*\step:\radius*1.43) {\footnotesize \em $\leq 6$ vertices in the \\ \footnotesize \em highlighted region};
\end{scope}
\end{tikzpicture}
\qquad \qquad \qquad
\begin{tikzpicture}[baseline=(current bounding box.north)]
  \begin{scope}
  \drawWLDfragment[7]{3}{0.4} 
  %\drawnumberspartial % use this to debug
  \centerarc[linehighlight](0,0)(\zero + 3.67*\step:\zero+4.43*\step:\radius)
  \newnode{1}{i}
  \newnode[below]{2}{i+1\ \ \ }
  \newnode[below]{3}{i+2}
  \newnode[below]{5}{j+1}
  \newnode[below]{6}{\quad j+2}
  \newnode[right]{7}{j+3}
      \begin{scope}
    \clipcenterarc(0,0)(\startpoint:\endpoint:\radius)
  \newpropbend{1.35}{3.6}{55}
  \newpropbend{4.4}{6.6}{55}
  \end{scope}
  \node[align=center] at (\zero + 4*\step:\radius*1.5) {\footnotesize \em $\leq 2$ vertices in the highlighted \\ \footnotesize \em region, and  no other propagators \\ \em \footnotesize end in the highlighted region};
\end{scope}
\end{tikzpicture}
\]
\caption{The two configurations described in Lemma~\ref{lem sian}; every weakly admissible Wilson loop diagram with no non-supporting vertices supports at least one example of at least one of these configurations.}\label{fig: small configs}
\end{figure}

@@

The second case is that there are at least $3$ vertices between one endpoint of $p_{r+1}$ and the nearest endpoint of $q_r$. In other words, (at least) one of the following four things happen:
\begin{itemize} 
\item $i_r+3 \leq j_{r+1}\leq k_r-3$ (one endpoint of $p_{r+1}$ has at least 3 vertices between it and the nearest enpoint of $q_r$), or
\item $i_r+3 \leq j_{r+1}+2\leq k_r-3$ (one endpoint of $p_{r+1}$ has at least 3 vertices between it and the nearest enpoint of $q_r$), or
\item $i_r = j_{r+1}$ ($p_{r+1}$ and $q_r$ share an end edge), or 
\item $k_r = j_{r+1} +2$  ($p_{r+1}$ and $q_r$ share an end edge)
\end{itemize}
In other words, at least one end of $p_{r+1}$ is supported entirely by vertices in the interval $[i_r+3,k_r+2]$ (the first two bullet points) or both ends lie in $[i_r, i_r+3]$ (third point) or both ends lie in $[k_r-2, k_r+1]$ (fourth point). These situations all behave similarly, and by symmetry it suffices to only consider the second and third possibilities.

>>>>

\begin{abstract}
  Wilson loop diagrams are an important tool in studying scattering amplitudes of SYM $N=4$ theory and are known by previous work to be associated to positroids. In this paper we study the structure of the associated positroids, as well as the structure of the denominator of the integral defined by each diagram. We give an algorithm to derive the Grassmann necklace of the associated positroid directly from the Wilson loop diagram, and a recursive proof that the dimension of these cells is thrice the number of propagators in the diagram. We also show that the denominator in the integral is the radical of the product of minors defined by the elements of the Grassmann necklace.
\end{abstract}
\section{Introduction}

@@

In recent years, there has been significant interest in understanding the geometry and combinatorics underlying the field theory SYM $N=4$ \cite{wilsonloop, Arkani-Hamed:2013jha, Amplituhedronsquared, galashinlam, AmplituhedronDecomposition}. This started with the observation that BCFW diagrams, which represent the on shell interactions of this theory, correspond to plabic graphs \cite{GrassmannAmplitudebook}. Plabic graphs were introduced by Postnikov in \cite{Postnikov}, and provide one of the many equivalent ways to classify the positroid cells of the positive real Grassmannian $\Gr(k,n)$. This led to the representation of the on shell integrals associated to this theory in terms of a geometric space (the Amplituhedron) embedded in $\Gr(k,n)$ \cite{Arkani-Hamed:2013kca,Arkani-Hamed:2013jha}. Moreover, integrals associated to the BCFW diagrams define a volume form on the Amplituhedron. This gives a novel and elegant interpretation of scattering amplitudes for this theory, as well as a means of greatly simplifying the calculations involved in traditional Feynman diagram approaches. Since then, there has been significant work to understand the volume form and its physical implications \cite{UnwindingAmplituhedron,galashinlam}.

Meanwhile, a different body of work has studied SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv,Adamo:2012xe,  Boels:2007qn, Bullimore:2010pj}. These integrals are calculated via holomorphic Wilson loops. As with the Amplituhedron, each Wilson loop represents the sum of a family of $N^kMHV$ Feynman diagrams, and adding up the corresponding sums of Feynman integrals associates a volume form to a convex space different from the Amplituhedron \cite{Hodges2013}. In  \cite{wilsonloop}, Agarwala and Marin-Amat uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that can be realized as an element of $\Gr(k,n)$. Other work has shown a connection between the volume filled by Wilson loop diagrams and the Amplituhedron \cite{casestudy, Amplituhedronsquared}. Specifically, Marcott shows that the space parametrized by the Wilson loop diagrams is  exactly the correct positroid cell in $\Gr(k,n)$, up to a set of measure $0$ \cite{WLDdim}.

@@

This paper continues this work by identifying which positroid cells are associated to which Wilson loop diagram: specifically, we give an algorithm to read off the Grassmann necklace of the associated positroid cell from the Wilson loop diagram. We also show that the positroid cells associated to Wilson loop diagrams are all $3k$-dimensional, where $k$ is the number of propagators in the diagram; by contrast, the BCFW cells are all $4k$-dimensional \cite{Arkani-Hamed:2013jha}. Again, this is not too alarming, as it was shown in \cite{non-orientable} that the actual geometric space defined by the Wilson loop diagrams is a non-orientable $k$-dimensional bundle over this $3k$ dimensional manifold in $\Gr(k,n)$.

@@

In Section~\ref{sec GN algorithm} we state an algorithm for constructing the Grassmann necklace of the positroid associated to a Wilson loop diagram (Algorithm~\ref{alg:put GN on WLD}) and prove its correctness (Theorem~\ref{res:alg gives GN}, Theorem~\ref{res alg gives correct GN}). While it was already possible to construct the Grassmann necklace of a Wilson loop diagram via existing bijections between some of the various combinatorial objects that index the positroid cells, this process was convoluted and involved multiple steps; our algorithm allows us to track directly how each propagator in the Wilson loop diagram contributes to each term in the Grassmann necklace. Furthermore, the above algorithm allows us to better understand the positroids associated to Wilson loop. In particular, Corollary \ref{no coloops} shows that these positroids have no coloops, and gives an exact characterization of their loops. Many of the technical lemmas required to prove Theorem~\ref{res:alg gives GN} and Theorem~\ref{res alg gives correct GN} continue to be important in subsequent sections: in particular we have Lemma~\ref{lem sian}, which identifies certain configurations of propagators that must appear in any admissible Wilson loop diagram, and Lemma~\ref{vertex cyclic int lem}, which shows that the pattern of values that a propagator contributes to each term of the Grassmann necklace is both simple and predictable.

@@

\subsection{Wilson loop diagrams}\label{section WLD background}

@@

A Wilson loop diagram is depicted as a circle, with the vertices $V$ arranged along the edge of the circle and listed counterclockwise in their cyclic order. The arc between two consecutive vertices is referred to as an {\bf edge} of the diagram, and the $i$th edge of $W$ is the edge from vertex $i$ to its successor. Each propagator $(i,j) \in \cP$ is denoted by a wavy line joining edge $i$ and edge $j$ inside the circle.

@@

We also impose the convention that all vertices should always be interpreted in terms of the cyclic order on $V$.  So, for any $i\in V$ we will use $i+1$ to denote the successor of $i$ in $V$.   In particular, if $V = [n]$ then $(n,n+3)$ means the propagator connecting edges $3$ and $n$. This viewpoint is beneficial for us since many of our proofs rely on restricting our attention to a region of the diagram bounded by a given propagator, or to short propagators such as $(i,i+2)$, and these behave the same regardless of whether we cross the ``first'' vertex or not. However, we emphasize that this is different to the standard convention in the physics literature, where propagators are written $(i,j)$ with $1 \leq i<j \leq n$.

@@

We will see in later sections that all four vertices supporting a propagator are important in understanding the contribution of that propagator to the objects represented by the diagram. This motivates the following definition of a subdiagram, which can consist of any subset $Q$ of the propagators and any subset $U$ of the vertices of $W$ but $U$ {\em must contain the support of $Q$}.

@@

\begin{dfn}\label{admisdfn}
A Wilson loop diagram $W = (\cP, V)$ is called {\bf admissible} if:
\begin{enumerate}
\item $|V| \geq |\cP| + 4$.
\item There does not exist a non-empty set of propagators $Q \subseteq \cP$ such that $|V(Q)| < |Q| + 3$.
\item There does not exist a pair of crossing propagators: that is, there does not exist ${\{(i,j),(k,l)\} \subseteq \cP}$ such that $i < k < j < l$.
\end{enumerate}
 \end{dfn}

@@

These diagrams correspond to (tree level) amplitudes of SYM $N=4$ theory in twistor space. As such, a diagram associates a real value to $n$ particles, one for each vertex. Each particle is represented as a section of a $|\cP|$-vector bundle over twistor space, projected onto a real subspace in a process called bosonization \cite{Arkani-Hamed:2013jha}. In other words, each Wilson loop diagram defines a functional on external particles. This is represented as an integral $I(W)$, which is the equivalent of a Feynman integral in this setting. This integral is comprised of two main components: an interaction matrix $C(W)$, and a volume form $\Omega(W)$. These are described next.

A $|\cP| \times n$ matrix $C(W)$ represents the (tree level) particle interactions. Each row of $C(W)$ corresponds to an internal propagator, and each column to an external particle (i.e. a vertex). $C(W)$ is defined as follows:

@@

The matrix $C(W)$ defines a positroid cell associated to $W$ in the positive Grassmannian $\Gr(n, |\cP|)$, which we denote by $\Sigma(W)$ \cite{wilsonloop}. Marcott \cite{WLDdim} shows that  the matrix $C(W)$ parametrizes a space whose \hlfix{close}{closure?} is exactly the closure of $\Sigma(W)$. That is, the space defined by $C(W)$ and the cell $\Sigma(W)$ agree up to a set of measure $0$. Most of the work in this paper and in the previous paper in the series \cite{generalcombinatoricsI} focuses on characterizing the positroid cell associated to each Wilson loop diagram $W$.

Note that since there is no ordering on the propagators of $W$, $C(W)$ is only defined up to rearrangement of the rows. When needed, we will impose various different orders on these rows to suit the task at hand. The choice of ordering does not affect $\Sigma(W)$.

@@

This implies that any singularities in the integrals $I(W)$ must cancel out in the sum; in particular, all poles arising from $R(W)$ must cancel. Specifically, these poles must cancel on the boundaries of the positroid cells $\Sigma(W)$, analogous to the analysis in \cite{Arkani-Hamed:2013jha}. This has been checked for particular classes of examples (e.g. \cite{casestudy, Amplituhedronsquared, HeslopStewart}) but has not been proven in general.

With this motivation in mind, in Section~\ref{sec poles} of this paper we study the denominator $R(W)$ of the volume form $\Omega(W)$. The definition of $R(W)$ is as follows.

\begin{dfn}\label{def R(W)}
Let $W = (\cP,V)$ be an admissible Wilson loop diagram, and fix an edge $e$. Let $\{p_1 \ldots p_r\}$ be the propagators with one end lying on edge $e$ under a counterclockwise ordering. That is, $p_1$ is closest to the vertex $e$, $p_i$ is positioned counterclockwise to $p_{i+1}$ and $p_r$ is the clockwise-most propagator on this edge (closest to vertex $e+1$). Define the polynomial
\[ R_e =  x_{p_1,e+1} \prod_{j= 1}^{r-1} \left((x_{p_j,e} x_{p_{j+1},e+1} - x_{p_{j+1},e} x_{p_{j},e+1} ) \right) x_{p_r,e}\;\]
to be the component of $R(W)$ associated to edge $e$. Note that if $r = 1$ (i.e. there is exactly one propagator lying on edge $e$) then this expression simplifies to $R_e = x_{p,e} x_{p,e+1}$. If $r=0$, set $R_e = 1$. The denominator $R(W)$ is defined to be the product of all of the $R_e$, i.e. 
\[R(W) = \prod_{e \in V} R_e.\]
\end{dfn}

In Section \ref{sec poles}, we show that the denominator $R(W)$ is determined by the positroid cell $\Sigma(W)$ and the structure of $W$, as encapsulated by $C(W)$. In fact, it can be read directly off of $C(W)$ in terms of a characterization of $\Sigma(W)$ called the Grassmann necklace (see Definition~\ref{def:grassmann necklace}). In Algorithm~\ref{alg WLD to denom via GN} we give an algorithm to read $R(W)$ directly from the Wilson loop diagram, paving the way for a general study of these poles in future.

@@

Let $W = (\cP,[n])$ be an admissible Wilson loop diagram with $k$ propagators, and $C(W)$ the associated matrix defined in equation~\eqref{C(W) dfn}. Let $M(W)$ be the matroid realized by $C(W)$, i.e. the matroid with ground set $[n]$ whose independent sets are exactly those sets $V \subseteq [n]$ such that the columns of $C(W)$ indexed by $V$ are linearly independent.

@@

  Alternatively we can simply think about the condition of linear independence in the matrix $C(W)$, and then the corollary can also be proved directly by linear algebra and the definition of $C(W)$ as follows.

  Because the nonzero entries of $C(W)$ are independent indeterminants, $J$ is an independent set if and only if there is some choice of $|J|$ nonzero entries of $C(W)$ one in each row associated to an element of $J$ and each in different columns. Each entry in $C(W)$ identifies a propagator by the row of the entry and a vertex by the column of the entry.  The entry is nonzero if and only if the propagator is supported on that vertex.  Consequently, a choice of $|J|$ nonzero entries of $C(W)$ one in each row associated to an element of $J$ and each in different columns is equivalent to an assignment of the propagators of $J$ to supporting vertices so that no two are assigned to the same vertex.  Such an assignment of the propagators of $J$ to supporting vertices is exactly a map $f$ as described in the statement, hence proving the result.
\end{proof}

@@

\begin{rmk}
While this paper primarily approaches proofs from a combinatorial point of view, readers with a background in matroid theory are encouraged to keep the matroidal viewpoint in mind throughout; we expect that the two approaches will be complementary and will build upon each other. See the first paper in this series \cite{generalcombinatoricsI} for some related results concerning the positroids of Wilson loop diagrams obtained by viewing them primarily as matroids.
\end{rmk}

\section{Wilson loop diagrams and their Grassmann necklaces}\label{sec GN algorithm}

In this section we will give an algorithm (Algorithm~\ref{alg:put GN on WLD}) to go directly from a Wilson loop diagram to the Grassmann necklace of its positroid.

This is a useful result in and of itself, as it greatly simplifies the process of identifying the positroid associated to a given Wilson loop diagram. We also use it to characterize loops and coloops in the positroid associated to a Wilson loop diagram (Corollary \ref{no coloops}); the characterization of the coloops, in particular, was not at all clear from the definition of Wilson loop diagrams but is an easy corollary of the algorithm.

Furthermore, the algorithm allows us to see how the Le diagram associated to a Wilson loop diagram changes as one adds propagators in certain positions (Theorem \ref{thm dim}). Indeed, the techniques we develop in this section to verify that our algorithm produces the correct Grassmann necklace will also be used repeatedly in later sections to calculate the dimension of the associated positroid cell (Section \ref{sec dim}) and to study the structure of the poles of the associated integral $I(W)$ (Section \ref{sec poles}).

One of the key insights is that each element of the Grassmann necklace can be viewed as a function from the propagators of the diagram to the vertices in the Grassmann necklace element.  This is captured in Definition~\ref{def I_i as a function}, and used throughout the paper.

@@

Before we can describe the algorithm for extracting the Grassmann necklace of $M(W)$ from the Wilson loop diagram $W$, we require some initial results about the behavior of propagators in admissible Wilson loop diagrams.

@@

Let $p$ be the propagator $p = (1, 4)$. Then $\cP_{in}(p, <_4) = \{(1,4), (5, 8), (5, 7) \}$ and $\cP_{out}(p, <_4) = \{(2,4) \}$
\end{eg}

@@

\begin{rmk}\label{rem:props of length 2 and 3} 
  The following observations about configurations of propagators of short length in a weakly admissibly Wilson loop diagram $W$ are easily verified:
 \begin{enumerate}
\item If $p = (i,i+3)$ is a propagator of length 3, then the middle vertex $i+2$ supports at most one propagator.
\item If every vertex in $W$ supports at least one propagator, then $W$ admits at least one propagator of length 2.  By the same reasoning, if $p=(i,j)$ is a propagator of $W$ and every vertex in the interval $[i,j]$ supports at least one propagator, then there is at least one propagator of length $2$ in $\mathcal{P}_{in}(p, <_i)$.
\end{enumerate}
\end{rmk}

\begin{figure}[h]
\[
\begin{tikzpicture}[baseline=(current bounding box.north)]
  \begin{scope}
  \drawWLDfragment[7]{3}{0.4} 
  %\drawnumberspartial % use this to debug
  \centerarc[linehighlight](0,0)(\zero + 1.43*\step:\zero+6.6*\step:\radius)
  \newnode{1}{}
  \newnode{3.55}{}
  \newnode{4.45}{}
  \newnode{7}{}
    \begin{scope}
    \clipcenterarc(0,0)(\startpoint:\endpoint:\radius)
  \newpropbend{1.3}{6.7}{20}
  \newpropbend{2.9}{5.1}{45}
  \end{scope}
  \node[align=center] at (\zero + 4*\step:\radius*1.43) {\footnotesize \em $\leq 6$ vertices in the \\ \footnotesize \em highlighted region};
\end{scope}
\end{tikzpicture}
\qquad \qquad \qquad
\begin{tikzpicture}[baseline=(current bounding box.north)]
  \begin{scope}
  \drawWLDfragment[7]{3}{0.4} 
  %\drawnumberspartial % use this to debug
  \centerarc[linehighlight](0,0)(\zero + 3.67*\step:\zero+4.43*\step:\radius)
  \newnode{1}{i}
  \newnode[below]{2}{i+1\ \ \ }
  \newnode[below]{3}{i+2}
  \newnode[below]{5}{j+1}
  \newnode[below]{6}{\quad j+2}
  \newnode[right]{7}{j+3}
      \begin{scope}
    \clipcenterarc(0,0)(\startpoint:\endpoint:\radius)
  \newpropbend{1.35}{3.6}{55}
  \newpropbend{4.4}{6.6}{55}
  \end{scope}
  \node[align=center] at (\zero + 4*\step:\radius*1.5) {\footnotesize \em $\leq 2$ vertices in the highlighted \\ \footnotesize \em region, and  no other propagators \\ \em \footnotesize end in the highlighted region};
\end{scope}
\end{tikzpicture}
\]
\caption{The two configurations described in Lemma~\ref{lem sian}; every weakly admissible Wilson loop diagram with no non-supporting vertices admits at least one example of at least one of these configurations.}\label{fig: small configs}
\end{figure}

@@

The second case is that there are at least $3$ vertices between one endpoint of $p_{r+1}$ and the nearest endpoint of $q_r$. In other words, at least one of the following four things happen:
\begin{itemize} 
\item $i_r+3 \leq j_{r+1}\leq k_r-3$ (one endpoint of $p_{r+1}$ has at least 3 vertices between it and the nearest endpoint of $q_r$), or
\item $i_r+3 \leq j_{r+1}+2\leq k_r-3$ (the other endpoint of $p_{r+1}$ has at least 3 vertices between it and the nearest endpoint of $q_r$), or
\item $i_r = j_{r+1}$ ($p_{r+1}$ and $q_r$ both have an end on edge $i_r$), or 
\item $k_r = j_{r+1} +2$  ($p_{r+1}$ and $q_r$ both have an end on edge $k_r$).
\end{itemize}
In other words, at least one end of $p_{r+1}$ is supported entirely by vertices in the interval $[i_r+3,k_r+2]$ (the first two bullet points) or both ends lie in $[i_r, i_r+3]$ (third point) or both ends lie in $[k_r-2, k_r+1]$ (fourth point). These situations all behave similarly, and by symmetry it suffices to only consider the second and third possibilities.

We have $j_{r+1}+4 \leq k_r-1$ in both of these cases: in the situation of the second bullet point this follows directly from the inequality, while in the third case it follows from the fact that $\ell(p_{r+1}) = 2$ and $\ell(q_r) \geq 4$. Therefore the vertex $j_{r+1}+4$ is in the interval $[i_r, k_r+1]$ but does not support either $q_r$ or $p_{r+1}$, and must therefore support some other propagator $t$.  Since $p_{r+1}$ is not part of configuration \ref{item pair of 2s} and $W_r$ has no propagators of length 3, it follows that $\ell(t) \geq 4$.  If $q_r$ and $p_{r+1}$ were on different sides of $t$ then this would contradict the minimality of $q_r$.  Therefore we can orient $t$ so that all previous $q_i$ and $p_i$ belong to $\cP_{out}(t)$, and set $q_{r+1} = t$ to continue the induction.

The overall result then follows by induction.
\end{proof}
